---
title: "p8105_hw5_kk3154"
author: "Kristen King"
date: "11/20/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  fig.width = 8, 
  fig.height = 6, 
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1 - Homicide Data

The Washington Post collected homicide data from 50 cities in the US for an [article](https://www.washingtonpost.com/graphics/2018/investigations/where-murders-go-unsolved/) and spatial analysis on unsolved murders. They shared their data in a [github repository](https://github.com/washingtonpost/data-homicides).

Problem 1 works with the homicide data to estimate the unsolved homicide rate in each city. 

Downloading the homicide data, summarizing the raw data, creating a combined city_state variable, and summarizing the total number of homicides and unsolved homicides (disposition of "Closed without arrest" or "Open/No arrest") by cities:
```{r}
raw_hom_data = read_csv(file = "./Data/homicide_data.csv")
skimr::skim(raw_hom_data)

homicide_df = read_csv(file = "./Data/homicide_data.csv", na = c("", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved"
    )) %>% 
  relocate(city_state) %>% 
  filter(city_state != "TulsaAL")

homicide_df %>% 
  group_by(city_state) %>% 
  summarize(homicide_count = n(),
            unsolved_homicides = sum(resolution == "unsolved")) %>% 
  arrange(homicide_count) %>% 
  knitr::kable(
    col.names = c("City", "Count of  Homicides", "Count of Unsolved Homicides")
  )
```

For the city of Baltimore, MD, use the `prop.test` function to estimate the proportion of homicides that are unsolved; save the output of `prop.test` as an R object, apply the `broom::tidy` to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "BaltimoreMD")
  
baltimore_summary = 
  baltimore_df %>% 
  summarize(
    unsolved = sum(resolution == "unsolved"),
    n = n()
  )

baltimore_test=  
  prop.test(
    x = baltimore_summary %>% pull(unsolved),
    n = baltimore_summary %>% pull(n))

baltimore_test %>% 
  broom::tidy()
```

Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of `purrr::map`, `purrr::map2`, list columns and `unnest` as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

Writing function and testing on a few cities:
```{r}
prop_test_function = function(city_df) {
  
  city_summary = 
    city_df %>% 
    summarize(
      unsolved = sum(resolution == "unsolved"),
      n = n())
  
  city_test = 
    prop.test(
      x = city_summary %>% pull(unsolved),
      n = city_summary %>% pull(n))
  
  return(city_test)

}

prop_test_function(baltimore_df)

homicide_df %>% 
  filter(city_state == "AlbuquerqueNM") %>% 
  prop_test_function()
```

Iterating across all cities:
```{r}
results_df = 
  homicide_df %>% 
  nest(data = uid:resolution) %>% 
  mutate(
    test_results = map(data, prop_test_function),
    tidy_results = map(test_results, broom::tidy)
  ) %>% 
  select(city_state, tidy_results) %>% 
  unnest(tidy_results) %>% 
  select(city_state, estimate, starts_with("conf"))

results_df
```

Alternate approach without creating a function, using map2:
```{r}
homicide_df %>% 
  group_by(city_state) %>% 
  summarize(homicide_count = n(),
            unsolved_homicides = sum(resolution == "unsolved")) %>% 
  mutate(
    test_results = map2(unsolved_homicides, homicide_count, prop.test),
    tidy_results = map(test_results, broom::tidy)
  ) %>% 
  select(city_state, tidy_results) %>% 
  unnest(tidy_results) %>% 
  select(city_state, estimate, starts_with("conf"))
```

Create a plot that shows the estimates and CIs for each city – check out `geom_errorbar` for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r}
results_df %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
















